---
title: "Causal inference for visual condition and optic flow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages 

```{r}
if(!require(xfun)) install.packages("xfun")
library(xfun)

if(!require(here)) install.packages("here")
library(here)

if(!require(data.table)) install.packages("data.table")
library(data.table)

if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if(!require(dplyr)) install.packages("dplyr")
library(dplyr)

if(!require(tidyr)) install.packages("tidyr")
library(tidyr)

if(!require(viridis)) install.packages("viridis")
library(viridis)

if(!require(rstan)) install.packages("rstan")
library(rstan)

if(!require(cmdstanr)) install.packages("cmdstanr")
library(cmdstanr)

if(!require(dagitty)) install.packages("dagitty")
library(dagitty)

devtools::install_github("rmcelreath/rethinking")
library(rethinking)
```

## Load response data 

- filter out participant 2 for now because visCond is not correct. Data needs reprocessing. 

```{r}
responses <- fread(file = here::here("regression_causal_inference/dat.responses.csv"))
```

## create visual condition variable

0 = no filter
1 = 1 filter
2 = 2 filter

```{r}
responses <- responses %>%
  dplyr::group_by(trialid) %>%
  dplyr::mutate(visCond.tmp = substr(trialid, 3, 4)) %>%
  dplyr::mutate(visCond = case_when(visCond.tmp == "_0" ~ 0,
                                    visCond.tmp == "_1" ~ 1,
                                    visCond.tmp == "_2" ~ 2))
```

## Creating a DAG

We have evidence that the day/night condition effect RTs. However we also know that altering the visual condition may effect their RT. The question is, is this a direct effect of visual condition, or does some of the effect flow through the day and night condition? 

This DAG implies that VC directly influences DN, VC directly influences RT, and that DN directly influences RT. Also that TTLC_f directly influences RT, but we can ignore this for now. 

```{r}
dag.1 <- dagitty("dag{VC -> DN; DN -> RT; VC -> RT; TTLC_f -> RT}")
coordinates(dag.1) <- list(x = c(VC = 0, RT = 1, DN = 2, TTLC_f = 3), y = c(VC = 0, RT = 1, DN = 0, TTLC_f = 1))
drawdag(dag.1)
```

## Plotting some data

```{r}
ggplot(responses %>%
         dplyr::filter(rt > 0, ttlc.failure == 4.12), mapping = aes(visCond, rt)) +
  geom_jitter()

ggplot(responses %>%
         dplyr::filter(rt > 0, ttlc.failure == 4.12), mapping = aes(dn, rt)) +
  geom_jitter()
```






